Measuring precise radial velocities and cross-correlation function line-profile variations using a Skew Normal density \thanks{Based on observations collected at the La Silla Parana Observatory, ESO (Chile), with the HARPS spectrograph at the 3.6-m telescope.}

### 2018-05-31 manuscript.tex
minor updates
abstract

\documentclass{aa}
%\documentclass[referee]{aa}
\usepackage{natbib}
%\usepackage{lscape}
\usepackage{pdflscape} % or {lscape}
\usepackage{wasysym}
\usepackage[varg]{txfonts}
\usepackage{graphicx}
\usepackage{hyperref}
\bibpunct{(}{)}{;}{a}{}{,} % to follow the A&A style
\usepackage{longtable}

@@

\author{U. Simola \inst{1,2}
	    \thanks{\email{umberto.simola@helsinki.fi}}
	    \and X. Dumusque\inst{3}
	    \thanks{Branco Weiss Fellow--Society in Science (url: \url{http://www.society-in-science.org})}    
	    \and Jessi Cisewski\inst{2}
	    }

@@

\abstract
% Context, Aims, Methods, Results, Conclu (not mandatory)
{Stellar activity is one of the primary limitations to the detection of low-mass exoplanets using the radial-velocity (RV) technique. 
Stellar activity can be probed by measuring time dependant variations in the shape of the cross-correlation function (CCF), often estimated using the different moments of the modelled CCF. Therefore estimating the moments of the CCF with high precision is essential to de-correlate exoplanet signals from spurious RV signals originating from stellar activity.}
%
{We propose to estimate the moments of the CCF by fitting a model using a Skew Normal (SN) density shape, which unlike the commonly employed Normal density, \xavier{includes an skewness parameter to capture the asymmetry of the CCF induced by stellar activity, but also the natural asymmetry induced by convective blueshift.}}
%
{The performance of the proposed method is compared to the Normal density using both simulated and real observations with varying levels of activity and signal-to-noise ratio (SNR) levels.}
%We analyze 5 stars with different activity levels and whose CCF's have different signal-to-noise ratio (SNR) levels. In each case, we compare the results obtained by fitting to the CCF respectively a Normal and a SN.  We also estimate rigorous errors for the different moments of the CCF using a bootstrap analysis.
%
{When considering the real observations, the correlation between the RV's and the asymmetry of the CCF and the correlation between the RV's and the width of the CCF are stronger when using the parameters derived from the SN than the Normal approach. 
This suggests that the CCF asymmetry and the CCF width derived using a SN may be more sensitive to stellar activity, which can be helpful with estimating stellar rotational periods and generally characterizing the stellar activity signals.
The estimated uncertainties in the estimated RV's using the proposed SN approach are on average $10\%$ smaller than the uncertainties calculated on the mean of the Normal, and the estimated uncertainties on the SN asymmetry parameter are on average $15\%$ smaller than the commonly used Bisector Inverse Slope Span (BIS SPAN) approach for estimating the asymmetry of the CCF. }
%
%{The derived benefits of using the proposed SN approach includes an asymmetry parameter along with several options for estimating the RV, while the Normal model only includes an RV estimate and requires a separate procedure for estimating the asymmetry.  Furthermore, the estimated uncertainties on the RV's and the asymmetry parameter are smaller in the SN setting.}
%The correlation between the RV's and the asymmetry of the CCF and the correlation between the RV's and the width of the CCF are always stronger when using the parameters derived from the SN in the case of real observations. Therefore the CCF asymmetry and the CCF width derived using a SN are more sensitive to stellar activity, which allows to probe with a better precision stellar rotational periods, in addition to characterize more precisely stellar activity signals. The precision on the estimated set of RV's, derived using the median of the fitted SN densities are on average $10\%$ smaller than the uncertainties calculated on the mean of the Normal. In addition, the uncertainties related to the asymmetry parameter derived from the SN are on average $15\%$ smaller than the ones calculated on the common Bisector Inverse Slope Span (BIS SPAN). We strongly encourage the use of the SN density rather than the Normal density, because this allows us to retrieve in one single fit the different moments of the CCF, because the derived moments better catch stellar activity signals and finally because the standard errors on the RV's and the the asymmetry parameter are smaller than the one estimated with the Normal fit.

>>>>

\documentclass{aa}
%\documentclass[referee]{aa}
%
\usepackage{pdflscape} % or {lscape}
\usepackage{natbib}
%\usepackage{lscape}
\usepackage{wasysym}
\usepackage[varg]{txfonts}
\usepackage{graphicx}
\usepackage{hyperref}
\bibpunct{(}{)}{;}{a}{}{,} % to follow the A&A style
\usepackage{longtable}

@@

\author{U. Simola \inst{1,2}
	    \thanks{\email{umberto.simola@helsinki.fi}}
	    \and X. Dumusque\inst{3}
	    \thanks{Branco Weiss Fellow--Society in Science (url: \url{http://www.society-in-science.org})}    
	    \and Jessi Cisewski-Kehe\inst{2}
	    }

@@

\abstract
% Context, Aims, Methods, Results, Conclu (not mandatory)
{Stellar activity is one of the primary limitations to the detection of low-mass exoplanets using the radial-velocity (RV) technique. 
Stellar activity can be probed by measuring time dependent variations in the shape of the cross-correlation function (CCF), often estimated using the different moments of the modeled CCF. Therefore estimating the moments of the CCF with high precision is essential to de-correlate exoplanet signals from spurious RV signals originating from stellar activity.}
%
{We propose to estimate the moments of the CCF by fitting a model using a Skew Normal (SN) density shape, which unlike the commonly employed Normal density, includes an skewness parameter to capture the asymmetry of the CCF induced by stellar activity, but also the natural asymmetry induced by convective blueshift.}
%
{The performance of the proposed method is compared to the Normal density using both simulated and real observations with varying levels of activity and signal-to-noise ratio (SNR) levels.}
%We analyze 5 stars with different activity levels and whose CCF's have different signal-to-noise ratio (SNR) levels. In each case, we compare the results obtained by fitting to the CCF respectively a Normal and a SN.  We also estimate rigorous errors for the different moments of the CCF using a bootstrap analysis.
%
{When considering the real observations, the correlation between the RV's and the asymmetry of the CCF and the correlation between the RV's and the width of the CCF are stronger when using the parameters derived from the SN than the Normal approach. 
This suggests that the CCF asymmetry and the CCF width derived using a SN may be more sensitive to stellar activity, which can be helpful with estimating stellar rotational periods and generally characterizing the stellar activity signals.
The estimated uncertainties in the estimated RV's using the proposed SN approach are on average $10\%$ smaller than the uncertainties calculated on the mean of the Normal, and the estimated uncertainties on the SN asymmetry parameter are on average $15\%$ smaller than the commonly used Bisector Inverse Slope Span (BIS SPAN) approach for estimating the asymmetry of the CCF. }
{XXX \jessi{It looks like we still need to add conclusions here}}
%
%{The derived benefits of using the proposed SN approach includes an asymmetry parameter along with several options for estimating the RV, while the Normal model only includes an RV estimate and requires a separate procedure for estimating the asymmetry.  Furthermore, the estimated uncertainties on the RV's and the asymmetry parameter are smaller in the SN setting.}
%The correlation between the RV's and the asymmetry of the CCF and the correlation between the RV's and the width of the CCF are always stronger when using the parameters derived from the SN in the case of real observations. Therefore the CCF asymmetry and the CCF width derived using a SN are more sensitive to stellar activity, which allows to probe with a better precision stellar rotational periods, in addition to characterize more precisely stellar activity signals. The precision on the estimated set of RV's, derived using the median of the fitted SN densities are on average $10\%$ smaller than the uncertainties calculated on the mean of the Normal. In addition, the uncertainties related to the asymmetry parameter derived from the SN are on average $15\%$ smaller than the ones calculated on the common Bisector Inverse Slope Span (BIS SPAN). We strongly encourage the use of the SN density rather than the Normal density, because this allows us to retrieve in one single fit the different moments of the CCF, because the derived moments better catch stellar activity signals and finally because the standard errors on the RV's and the the asymmetry parameter are smaller than the one estimated with the Normal fit.
